A general model of microbial interactions in a chemostat

### 2019-12-02 tex/Tracking.tex
Update tex

\begin{abstract}
	This article deals with the inclusion of microbial ecology in bioprocess modelling. The first part presents the mathematical analysis of a model that may be framed within the class of Lotka-Volterra models fitted to experimental data in a chemostat setting. The limitations and the insights of such an approach are discussed. The main novelty of this work lies in the second part, where the use of an optimal tracking technique for the integration of genetic sequencing in chemostat models is presented. The method revisits the data used for the first model presented and proposes a way to calibrate and discriminate for a class . 
\end{abstract}

@@

The main contribution of this article is a procedure to integrate the microbial community measurements in chemostat models, based on any technique that can quantify the species abundances through time. This is a first step towards developing predictive models of a reactor's performance based on the microbial community structure.

@@

Section 4 is the article's most original contribution. The concept of interaction function is developed such that it generalizes the gLV model. For the identification of the interaction function a method of optimal tracking was adapted: The growth rate of each species is modulated by a constrained regular control of the system, thus the growth rate of each OTU is corrected to fit the experimental data. The regular control is composed of a feedback part on the species state variable, and a feed forward part, or tracking, on the measurement of abundances of each species.

In section 5 the methodology from section 4 is applied to the data from the experiments perform by Dumont et al \cite{Dumont2009}. Finally, a discussion on the scope of the method, its shortcomings and perspectives is presented.

@@

The washout of $G_2$ is equivalent to the following hypothesis:

@@

The model here presented serves to illustrate that by considering a more complex growth rate that accounts for ecological interactions one might explain differences in reactors operating under similar conditions. It also shows a new mechanism by which the CEP no longer holds. However the quadratic increase of parameters with the number of OTU considered makes it difficult to accurately calibrate the model. The operating diagrams always show a single stable equilibrium point for each pair $(s_{in},D)$, however no proof is offered to such conjecture. In the next section, a generalization of the former model, and a schema for calibrating a class of models is presented as well.

@@

Finally for the parametric identification of a particular class of interaction functions an static optimization problem is proposed \eqref{Parameter Identification}. Suppose $u(t)$ is the optimal control of problem \eqref{Optimal Control}, and $x(t)$ is the associated optimal trajectory. Suppose a set of parameters $\theta \in \Theta$, such that the interaction function has a explicit form that depends on this parameters $I_{\theta}(x)$ ($\theta$ would represent matrix $A$ from section 2 for example). 
\begin{align}
\label{Parameter Identification} \min & \quad \Vert I_{\theta}(x(\cdot)) - u(\cdot) \Vert_{L^2([0,T])} \\
s.t. &  \quad \theta \in \Theta \notag \\
& \quad t \in [0,T] \notag
\end{align}

In practice $t \in [0,T]$ from \eqref{Parameter Identification} is replaced by considering the time steps given by the differential equation solver.
\subsection{Proof of concept}

For compensating the lack of a theoretical optimal control, the approach presented in here was first tested against synthetic data. The test consisted in simulating system \eqref{system} with growth functions \eqref{gLV growth1} and \eqref{gLV growth2} (therefore $I(x) = 1 + Ax$) for $n_1 = 10$ and $n_2 = 5$. In all figures, dotted lines represent the synthetic data, while the continuous lines represent the method's output.

@@

The method stopped after 110 iterations where each iteration took approximately 600 seconds of computing time in a computer equipped with 8gb of RAM memory and Intel core i3-7100U CPU 2,40 GHz.

The tracking problem was applied to data coming from a nitrification process; the experimental conditions are described in \cite{dumont2008observers}. For exploring the hypothesis of interactions as drivers of bioreactors performance environmental conditions should be kept as constant as possible. Therefore only data from day $183$ onwards was used because a change in the operating temperature happened at that point, which is known to have an effect on the kinetics.  For choosing which species belong in which functional group, the procedure described in \cite{Ugalde-Salas2019} was used: from day $183$ to day $230$: 17 OTU were identified in the $G_1$ group (known as Ammonium oxidizing bacteria AOB) and 8 in the $G_2$ functional group (known as Nitrite oxidizing bacteria NOB). The system state dimension becomes 28.

@@

One can observe that in figure \ref{ControlSimulation} the predicted $s_3$ around day 240 is much higher than the measurement. If one looks at figure \ref{NOB Simulation} it suggests attributing this extra production of $s_3$ around day 240 to the growth of OTU 24. What would have happened if OTU 24 had been classified in the other functional group? One would have seen a decrease in $s_1$, an increase in $s_2$, and a decrease in $s_3$ which could have better fitted the measurements. Thus suggesting that OTU 24 was misclassified from the method presented in \cite{Ugalde-Salas2019}.

@@

Recalling that $u(t)$ is supposed to represent $I(x(t))$ (the interactions terms), one could try to fit different forms of $I(\cdot)$ for example $I(x) = 1 + Ax$, where  $A$ is a real valued square matrix. In figure \ref{LV_fit} the simulation of the system after the identification of matrix A is presented, suggesting that an affine interaction function is not able to capture the behaviour of the system.

@@

The interaction function introduced tries to address the issue of prediction from the genetic sequencing, leaving the conclusion that linear pairwise interaction model do not capture the complexity of the particular nitrifying chemostat analysed. Furthermore from the proof of concept section, one might conjecture that models with a large number of parameters such as an gLV model may be unidentifiable, thus proving to be a real problem in calibration, and therefore of reduced applicability. This opens the question for a broader class of interaction functions that could fit well complex microbial ecosystems. The experimental study in chemostat systems seems a good compromise between completely open ecosystems (nature) and controlled conditions (laboratory) in order to explore expressions that could account for microbial interactions.

@@

\section{Acknowledgements}
 Alain Rapaport, CMM, Thermomic

@@

They were solved using standard backward numerical integration.

\bibliographystyle{elsarticle-num}
\bibliography{mybibfile}
\end{document}

>>>>

\begin{abstract}
	This article deals with the inclusion of microbial ecology advances in bioprocess modelling. The first part presents the mathematical analysis of a model that may be framed within the class of Lotka-Volterra models fitted to experimental data in a chemostat setting. The limitations and the insights of such an approach are discussed. The main novelty of this work lies in the second part, where the use of an optimal tracking technique for the integration of genetic sequencing in chemostat models is presented. The method revisits the data used for the first model presented and proposes a way to calibrate and discriminate for a class of growth kinetics. 
\end{abstract}

@@

The main contribution of this article is a procedure to integrate the microbial community measurements in chemostat models, based on any sequencing technique that can quantify the species abundances through time. This is a first step towards developing predictive models of a reactor's performance based on the microbial community structure.

@@

Section 4 is the article's core contribution. The concept of interaction function is developed such that it generalizes the gLV model. For the identification of the interaction function a method of optimal control theory was adapted: The growth rate of each species is modulated by a constrained regular control of the system, thus the growth rate of each OTU is corrected in order to fit the experimental data. The regular control is composed of a feedback part on the species state variable, and a feed forward part, or tracking, on the measurement of abundances of each species, it is solved via state-dependent ricatti equations \cite{Cimen2008}. Finally the obtained optimal trajectory and control are used in a static optimization problem in order to estimate the parameters of a specific interaction function.

In section 5 the methodology from section 4 is applied to all of the data from the experiments performed by Dumont \textit{et al}\cite{Dumont2009} and not just to the most abundant species as it was the case of \cite{Dumont2016}. The article ends with a discussion on the use of gLV models, the scope of applicability of the method, and its shortcomings.

@@

The washout of $G_2$ is equivalent to  hypothesis \ref{WashoutG2}.

@@

The model here presented serves to illustrate that by considering a more complex growth rate that tries to model ecological interactions one might explain differences in reactors operating under similar conditions. It also shows a new mechanism by which the CEP no longer holds. The operating diagrams always show a single stable equilibrium point for each pair $(s_{in},D)$, however no proof is offered to such conjecture.
The quadratic increase of parameters with the number of OTU considered makes it difficult to accurately calibrate the model. Parameter estimation in A classic way to approach this is by a least square approach, however Some methods for parameter estimation and uncertainty in biochemical systems of ODE can be seen in Vanlier \textit{et al.} \cite{Vanlier2013} and even a "robust approach" is proposed for biological systems by Gabor \textit{et al.} \cite{Gabor2015}, however parameter estimation in ODE is still an open problem. In the next section, a generalization of the former model, and a schema for parameter estimation where the uncertainty lies in the growth function is presented as well.

@@

Finally for the parametric identification of a particular class of interaction functions a static optimization problem is proposed \eqref{Parameter Identification}. Suppose $u(t)$ is the optimal control of problem \eqref{Optimal Control}, and $x(t)$ is the associated optimal trajectory. Suppose a set of parameters $\theta \in \Theta$, such that the interaction function has a explicit form that depends on this parameters $I_{\theta}(x)$ ($\theta$ would represent matrix $A$ from section 2 for example). In practice $t \in [0,T]$ from \eqref{Parameter Identification} is replaced by considering the time steps given by the differential equation solver. 
\begin{align}
\label{Parameter Identification} \min & \quad \Vert I_{\theta}(x(\cdot)) - u(\cdot) \Vert_{L^2([0,T])} \\
s.t. &  \quad \theta \in \Theta \notag \\
& \quad t \in [0,T] \notag
\end{align}

\subsection{Proof of concept}

The approach was tested against synthetic data. The test consisted in simulating system \eqref{system} with growth functions \eqref{gLV growth1} and \eqref{gLV growth2} (therefore $I(x) = 1 + Ax$) for $n_1 = 10$ and $n_2 = 5$. In all figures, dotted lines represent the synthetic data, while the continuous lines represent the method's output.

@@

The method was stopped after 10 iterations where each iteration took approximately 800 seconds of computing time in a computer equipped with 8gb of RAM memory and Intel core i3-7100U CPU 2,40 GHz.

The tracking problem was applied to data coming from a nitrification process; the experimental conditions are described in \cite{dumont2008observers}. For exploring the hypothesis of interactions as drivers of bioreactors performance environmental conditions should be kept as constant as possible. Therefore only data from day $183$ onwards was used because a change in the operating temperature happened at that point, which is known to have an effect on the kinetics. For choosing which species belong in which functional group, the procedure described in \cite{Ugalde-Salas2019} was used: from day $183$ to day $315$: 30 OTU were identified in the $G_1$ group (known as Ammonium oxidizing bacteria AOB) and 5 in the $G_2$ functional group (known as Nitrite oxidizing bacteria NOB). The system state dimension becomes 38.

@@

One can observe that in figure \ref{ControlSimulation} the predicted $s_3$ around day 240 is much higher than the measurement.

@@

Then the procedure from the proof of concept section was applied and one can see that a gLV model is far from describing the pronounced control changes.

@@

The interaction function introduced tries to address the issue of prediction from the genetic sequencing, leaving the conclusion that linear pairwise interaction model do not capture the complexity of the particular nitrifying chemostat analysed. Furthermore from the proof of concept section, one might conjecture that models with a large number of parameters such as an gLV model may be unidentifiable, thus proving to be a real problem in calibration, and therefore of reduced applicability. Furthermore the ecological interpretations to be drawn from the sign of the interaction matrix might be useless. This opens the question for a broader class of interaction functions that could fit well complex microbial ecosystems. The experimental study in chemostat systems seems a good compromise between completely open ecosystems (nature) and controlled conditions (laboratory) in order to explore expressions that could account for microbial interactions.

@@

\section{Codes}

All the codes can be recovered from the following repository \url{https://github.com/paus-5/Class-and-Track}.
\section{Acknowledgements}
 The authors would like to thank Alain Rapaport for the fruitful exchanges. This work was financed by grant Thermomic. And a sejour

@@

They were solved using standard backward numerical integration.
\section{References}
\bibliographystyle{elsarticle-num}
\bibliography{mybibfile}
\end{document}
